Detecting Manual Alterations in Biological Image Data Using Contrastive Learning and Pairwise Image Comparison

### 2025-03-20 paper/main.tex
Computational Experiment + Code

\usepackage{subcaption}
\usepackage{graphicx}
\usepackage{multirow}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{amsthm}
\usepackage{mathrsfs}
\usepackage{xcolor}
\usepackage{textcomp}
\usepackage{manyfoot}
\usepackage{booktabs}
\usepackage{algorithm}
\usepackage{algorithmicx}
\usepackage{algpseudocode}
\usepackage{listings}

@@

>>>>

\usepackage{subcaption}
\usepackage{graphicx}
\usepackage{multirow}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{amsthm}
\usepackage{mathrsfs}
\usepackage{xcolor}
\usepackage{textcomp}
\usepackage{manyfoot}
\usepackage{booktabs}
\usepackage{algorithm}
\usepackage{algorithmicx}
\usepackage{algpseudocode}
\usepackage{listings}
\usepackage{caption}

\captionsetup[figure]{width=0.7\textwidth}

@@

\section{Computational Experiment}

In this section, we conduct an experiment to train the projector of the 
\texttt{Barlow Twins} model on a blood cells dataset. 
Our goal is to determine if the model, with its state-of-the-art pretrained encoder, 
can be successfully adapted to our problem.

The model's encoder is the pretrained \texttt{ResNet50} from the original article
\cite{zbontar2021barlow}. The dataset consists of $320$ images of lymphocyte cells, 
split into a training set of $256$ images and a validation set of $64$ scans. 
The projector is composed of three groups of layers; each group contains a linear layer, 
batch normalization, and a ReLU activation function. 
The input dimensions for the groups are $2048$, $1024$, and $1024$, respectively.

The loss function is identical to that described in the original article. 
For optimization, we use the \texttt{Adam} optimizer with the following learning rate schedule:

$$ q_k = \frac{1}{2} \cdot (1 + \cos(\pi \cdot \frac{k}{K})) $$

$$ \gamma_{k} = \gamma_{start} \cdot q_k + \gamma_{end} \cdot (1 - q_k) $$

where $k$ is the epoch number, $K$ is the total number of epochs, 
$\gamma_{start} = 5 \cdot 10^{-3}$, $\gamma_{end} = 5 \cdot 10^{-4}$.

\begin{figure}[htbp]
    \centering
    \includegraphics[width=0.95\textwidth]{figures/BT_blood-cells-256_experiment.png}
    \caption{Graph of the loss function for training and validation samples. 
    The experiment was conducted over $180$ epochs. 
    The validation loss decreases alongside the training loss, indicating that 
    the model can be trained further. 
    Training takes approximately $25$ minutes on a Google Colab GPU. 
    }
    \label{fig:experiment}
\end{figure}
